---
title: 'The P600 for singular ''they'' v2:  P600 analyses for Centro-Parietal Electrode Sites'
author: "Joanna Morris"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
fontsize: 10pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, fig.dim = c(4, 3))
library(knitr)
library(ggplot2)
library(ggrepel)
library(psych)
library(readr)
library(stringr)
library(dplyr)
library(tidyr)
library(pander)
library(wesanderson)
library(ez)

```


This document contains the code to format the data required to conduct the statistical analyses for PROST project. It takes participants with more than 65% good data from Grusha's original datasets (2017), label all the variables, collapses over all the channels  and then creates a difference wave. The output files  are "prost_2022_singular_p600.csv" and "prost_2022_plural_p600.csv" 


### Loading in and formatting the data

(1) First we get the list of participants with more than 65% good data from Grusha's original datasets (2017)

```{r}
group2 <- read.table('group2_less_than_35_percent_data.txt', header = TRUE)
group2$Group = "Binary"

group3 <- read.table('group3and4_less_than_35_percent_data.txt', header = TRUE)
group3$Group = "Non-binary"

all_data = rbind(group2, group3)

less_than_35_percent_subjs <- unique(all_data$ERPset)
write.csv(less_than_35_percent_subjs
          , "prost_subjlist_lessthan35.csv"
          , row.names = FALSE)

rm(all_data)
rm(group2)
rm(group3)

```

(1) Then we load in the data and label all the variables. 

```{r}

prost_2022 <- read_csv('prost_mea_500800_202207.csv')


centroparietal_channels  <- c('C3','Cz', 'C4', 'CP3', 'CPz', 'CP4', 'P3',  'Pz',  'P4')

posterior_channels <- c('PZ','P4', 'P8', 'CP4', 'P7',  'P3',  'CP3', 'CPZ', 'TP7', 'TP8')

nref_channels <- c('F3', 'Fz', 'F4', 'FC3', 'FCz', 'FC4', 'C3','Cz', 'C4', 'CP3', 'CPz', 'CP4', 'P3',  'Pz',  'P4')

prost_2022 <- dplyr::filter(prost_2022, ERPset %in% less_than_35_percent_subjs)
prost_2022 <- dplyr::filter(prost_2022, chlabel %in% posterior_channels )


prost_2022 <- dplyr::mutate(prost_2022,
                            SubjID = str_extract(prost_2022$ERPset, "\\d{3}"))

prost_2022 <- dplyr::mutate(prost_2022, 
                            Referentiality = case_when(grepl("NonReferential"
                                                             ,binlabel) ~ "NonReferential", 
                                                       grepl("Referential"
                                                             , binlabel) ~ "Referential"))

prost_2022 <- dplyr::mutate(prost_2022, 
                            Gender_Status = case_when(grepl("NonGendered"
                                                            ,binlabel) ~ "NonGendered", 
                                                       grepl("Gendered"
                                                             , binlabel) ~ "Gendered"))

prost_2022 <- dplyr::mutate(prost_2022, 
                            Critical = case_when(grepl("Critical"
                                                       ,binlabel) ~ "Critical",
                                                 grepl("Baseline"
                                                       , binlabel) ~ "Baseline"))

prost_2022 <- dplyr::mutate(prost_2022, 
                            Group = case_when(grepl("prost4"
                                                    ,ERPset) ~ "NonBinary",
                                              grepl("prost3"
                                                    , ERPset) ~ "NonBinary",
                                              grepl("prost2"
                                                    , ERPset) ~ "Binary"))

prost_2022 <-dplyr::mutate(prost_2022, 
                            Number = case_when(grepl("himself_herself"
                                                     ,binlabel) ~ "Singular",
                                               grepl("themselves"
                                                     , binlabel) ~ "Plural"))

```


(3) Then we collapse over the channels and gets difference scores

```{r}
# Collapse over all channels
library(doBy)
prost_2022_2<- 
  summaryBy(value ~ SubjID + Referentiality + Gender_Status + Critical + Group + Number
            , data=prost_2022)


# get difference scores (Critical - Baseline)
prost_2022_3 <- 
  pivot_wider(prost_2022_2
              , names_from = Critical
              , values_from = value.mean)

prost_2022_3 <- 
  mutate(prost_2022_3
         , diff_score = Critical - Baseline)
```


(4) Then we convert character vectors to factors

```{r, results=FALSE}
names <- c('Referentiality' ,'Gender_Status', 'Group', 'Number')
prost_2022_3[,names] <- lapply(prost_2022_3[,names] , factor)
str(prost_2022_3)
```

(5) Then we re-order factor levels for *Referentiality*

```{r}
prost_2022_3$Referentiality <- factor(prost_2022_3$Referentiality, 
                                            levels=c('Referential',
                                                     'NonReferential'))
levels(prost_2022_3$Referentiality)
```


```{r}
prost_2022_singular <- dplyr::filter(prost_2022_3, Number == "Singular") 
prost_2022_singular <- select( prost_2022_singular, -Number)
write_csv(prost_2022_singular, "prost_2022_singular_p600.csv")

prost_2022_plural <- dplyr::filter(prost_2022_3, Number == "Plural")
prost_2022_plural <- select( prost_2022_plural, -Number)
write_csv(prost_2022_plural, "prost_2022_plural_p600.csv")

```


