
---
title: "The P600 effect when singular gendered antecedents are co-indexed with (a) *himself* or *herself* (b) *themselves*"
author: "Joanna Morris"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

This script, on the advice of reviewer 1, conducts an ANOVA examining the P600 PROST data with Referentiality and Gender as within-subject variables. Gender Identity status will be examined as a post-hoc variable. Analysis 1 examines the data for antecedents that are co-indexed with *himself* or *herself*, while Analysis 2 examines the  P600 effect when antecedents are co-indexed with *themselves*



```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, error = FALSE, message = FALSE, fig.dim = c(6, 6))
library(knitr)
library(ggplot2)
library(ggrepel)
library(psych)
library(readr)
library(dplyr)
library(tidyr)
library(pander)
library(ggsci)
library(ez)
library(gridExtra)
```

\scriptsize

## Define functions, set parameters and load

Define standard error of mean function

```{r echo=TRUE}
sem <- function(x) sd(x)/sqrt(length(x))
```

Before we begin, let's set some general parameters for `ggplot2`. We will set a general theme using the `theme_set()` function. We will use the 'classic' theme which gives us clean white background rather than the default grey with white grid lines. We will position the legend at the top of the graph rather than at the right side which is the default.

```{r theme}
theme_set(theme_classic()+
            theme(legend.position = "bottom")) 
```


```{r}
prost_2022_singular <- read_csv("prost_2022_singular_p600.csv")
prost_2022_plural <- read_csv("prost_2022_plural_p600.csv")
prost_2024_combined <- bind_rows(list("Gendered Singular" = prost_2022_singular, 
                                      "NonGendered Plural" = prost_2022_plural), .id = "Anaphor_Type")
```

Then we re-order factor levels for  *Referentiality* and *Anaphor_Type*

```{r results='hide'}
prost_2024_combined$Referentiality <- factor(prost_2024_combined$Referentiality, 
                                            levels=c('Referential',
                                                     'NonReferential'))
prost_2024_combined$Anaphor_Type <- factor(prost_2024_combined$Anaphor_Type, 
                                            levels=c('Gendered Singular',
                                                     'NonGendered Plural'))
levels(prost_2024_combined$Referentiality)
levels(prost_2024_combined$Anaphor_Type)
```

## Analysis 

```{r echo=TRUE}

pander(ezANOVA(data = prost_2024_combined
              , dv = diff_score
              , wid = SubjID
              , within = .(Referentiality, Gender_Status, Anaphor_Type)
              , type = 3
              , return_aov = F
              ))
```

### Condition Means for Analysis 1


Significant Effects: **Referentiality; Referentiality x Anaphor Type; Referentiality X Gender Status X Anaphor Type**

```{r}

kable(singular_means1 <- prost_2024_combined |> 
  group_by(Referentiality) |>
  summarise(Mean = mean(diff_score), 
            SE = sem(diff_score), 
            SD = sd(diff_score), 
            Max = max(diff_score), 
            Min = min(diff_score)), digits = 2)

kable(singular_means1 <- prost_2024_combined |> 
  group_by(Referentiality, Anaphor_Type) |>
  summarise(Mean = mean(diff_score), 
            SE = sem(diff_score), 
            SD = sd(diff_score), 
            Max = max(diff_score), 
            Min = min(diff_score)), digits = 2)


kable(singular_means2 <- prost_2024_combined |> 
  group_by( Anaphor_Type,  Gender_Status, Referentiality) |>
  summarise(Mean = mean(diff_score), 
            SE = sem(diff_score), 
            SD = sd(diff_score), 
            Max = max(diff_score), 
            Min = min(diff_score)), digits = 2)

```

### Post-hoc tests 

Runs post-hoc tests for the 3-way ***"Gender Status x Referentiality * Anaphor Type"*** Interaction


"Some woman...himself" vs. "Mary...himself"  

```{r}
pander(t.test(diff_score ~ Referentiality
       , filter(prost_2024_combined, Anaphor_Type == "Gendered Singular" & Gender_Status == "Gendered" )))
```

"Someone...himself" vs. "The participant...himself" 

```{r}
pander (t.test(diff_score ~ Referentiality
       , filter(prost_2024_combined, Anaphor_Type == "Gendered Singular" & Gender_Status == "NonGendered")))
```

"Some woman...themselves" vs. "Mary..themselves"  

```{r}
pander(t.test(diff_score ~ Gender_Status
       , filter(prost_2024_combined, Anaphor_Type == "Gendered Plural" &  Gender_Status == "Gendered")))
```

"Someone...themselves" vs. "The participant...themselves"  
 
```{r}
pander(t.test(diff_score ~ Gender_Status
       , filter(prost_2024_combined, Anaphor_Type == "Gendered Plural" & Gender_Status == "NonGendered" )))
```

##### Interaction Plots
&nbsp;

```{r}
summary_data_2 <- prost_2024_combined |>
  group_by(Anaphor_Type,Gender_Status, Referentiality) |>
  summarise(mean = mean(diff_score), se = sem(diff_score))|> 
  mutate(lwr = mean - (1.96*se), upr = mean + (1.96*se))
summary_data_2$label = c("John...herself","some man...herself","the participant...herself", "someone...herself", "John...themselves", "some man...themselves", "the participant...themselves", "someone...themselves" )

(plotA <- ggplot(summary_data_2, aes(x = Gender_Status, 
                         y = mean,
                         ymin = lwr,
                         ymax = upr,
                         fill = Referentiality,
                         colour = Referentiality,
                         label = label)) + 
    geom_bar(position=position_dodge(), stat="identity") + 
    geom_errorbar(width = .15, position = position_dodge(.9), colour = "black")  +
    geom_label_repel(show.legend = FALSE, cex = 1.5, colour = "black") +
    ylab("Voltage in microvolts") +
    theme(legend.text = element_text(size=10))+ 
    theme(legend.title = element_blank()) +
    facet_wrap(~ Anaphor_Type, ncol = 2) +
    scale_color_npg() +
    scale_fill_npg(alpha = 0.8))
```






