---
title: "The N400 effect when singular gendered antecedents are co-indexed with  *himself* or *herself*" 
author: "Joanna Morris"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

\scriptsize

## Overview

This document contains the code to reproduce the statistical analyses described in [Prasad and Morris (2024)]

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, error = FALSE, message = FALSE, fig.dim = c(7, 5 ))
library(knitr)
library(ggplot2)
library(ggrepel)
library(psych)
library(readr)
library(dplyr)
library(tidyr)
library(pander)
library(ez)
library(ggsci)

```



## Define functions, set parameters and load

Define standard error of mean function

```{r echo=TRUE}
sem <- function(x) sd(x)/sqrt(length(x))
```

Before we begin, let's set some general parameters for `ggplot2`. We will set a general theme using the `theme_set()` function. We will use the 'classic' theme which gives us clean white background rather than the default grey with white grid lines. And we will position the legend at the top of the graph rather than at the right side which is the default.

```{r theme}
theme_set(theme_classic() +  theme(legend.position = "bottom")) 
```

## Load the data

```{r}
prost_2024_singular <- read_csv("prost_2022_singular_n400_nref.csv")
prost_2024_plural <- read_csv("prost_2022_plural_n400_nref.csv")
```

Then we re-order factor levels for *Anteriority* & *Referentiality*

```{r}
prost_2024_singular$Anteriority <- factor(prost_2024_singular$Anteriority, 
                                            levels=c('Frontal',
                                                     'FrontoCentral', 
                                                     'Central', 
                                                     'CentroParietal', 
                                                     'Parietal'))

prost_2024_singular$Referentiality <- factor(prost_2024_singular$Referentiality, 
                                            levels=c('Referential',
                                                     'NonReferential'))
levels(prost_2024_singular$Anteriority)
levels(prost_2024_singular$Referentiality)

prost_2024_plural$Anteriority <- factor(prost_2024_plural$Anteriority, 
                                            levels=c('Frontal',
                                                     'FrontoCentral', 
                                                     'Central', 
                                                     'CentroParietal', 
                                                     'Parietal'))

prost_2024_plural$Referentiality <- factor(prost_2024_plural$Referentiality, 
                                            levels=c('Referential',
                                                     'NonReferential'))
levels(prost_2024_plural$Anteriority)
levels(prost_2024_plural$Referentiality)
```

## Check ANOVA assumptions

- *No significant outliers in any cell of the design*. This can be checked by visualizing the data using box plot methods and by using the function `identify_outliers()` in the `rstatix` package.

```{r}
library(rstatix)
kable(identify_outliers(prost_2024_singular, diff_score))
```

- *Normality*: the outcome (or dependent) variable should be approximately normally distributed in each cell of the design. This can be checked using the Shapiro-Wilk normality test `shapiro_test()`in the `rstatix` package.

```{r}
kable(shapiro_test(prost_2024_singular, diff_score))
```

- *Assumption of sphericity*: the variance of the differences between groups should be equal. This can be checked using the Mauchly’s test of sphericity, which is automatically reported when using the R function `anova_test()` in the `rstatix` package.


## Analysis using `rstatix()`" The N400 effect when antecedents are co-indexed with *himself* or *herself*


```{r}
library(kableExtra)
fitted.model.0 <- anova_test(data = prost_2024_singular, 
                      dv = diff_score, 
                      wid = SubjID, 
                      within = c(Referentiality, Gender_Status, Anteriority),
                      between = Group)
kable(fitted.model.0$ANOVA)
kable(fitted.model.0$`Mauchly's Test for Sphericity`)
kable(fitted.model.0$`Sphericity Corrections`)
```

## Analysis using `ezANOVA`: The N400 effect when antecedents are co-indexed with *himself* or *herself*


```{r echo=TRUE}

fitted.model.1 <- ezANOVA(data = prost_2024_singular
              , dv = diff_score
              , wid = SubjID
              , within = .(Referentiality, Gender_Status, Anteriority)
              , between = Group
              , type = 3
              , return_aov = F)
kable(fitted.model.1$ANOVA)
kable(fitted.model.1$`Mauchly's Test for Sphericity`)
kable(fitted.model.1$`Sphericity Corrections`)
```


## Analysis using lmer

From [Introduction to Regression Methods for Public Health Using R by Ramzi W. Nahhas](https://www.bookdown.org/rwnahhas/RMPH/mlr-distinctions.html#)
Use `car::Anova(, type = 3)` to test whether all the adjusted regression coefficients associated with a single term in the model (e.g., a single predictor, a single interaction) are simultaneously zero. For binary categorical predictors, continuous predictors, and interactions between them, the `car::Anova()` output is redundant with the regression coefficient table output from summary() since each such term only has one corresponding regression coefficient. However, you must use `car::Anova()` to test the significance of a categorical predictor with more than two levels or of an interaction that involves a categorical predictor with more than two levels. `car::Anova()` carries out comparisons for each of a set of specific pairs of nested models – each comparison is between the full model and a reduced model with one predictor (or interaction) removed.

```{r echo=TRUE, warning=FALSE, message=FALSE}
library(lme4)
library(car)
fitted.model.2 <- lmer(diff_score ~ Referentiality * Gender_Status * Anteriority * Group + (1|SubjID), data=prost_2024_singular)
kable(Anova(fitted.model.2, type = 3))
```



### Post-hoc tests : Group x Gender Status x Referentiality

If there is a significant three-way interaction effect, you can decompose it into:

- *Simple two-way interaction*: run two-way interaction at each level of third variable,

- *Simple simple main effect*: run one-way model at each level of second variable, and

- *Simple simple pairwise comparisons*: run pairwise or other post-hoc comparisons if necessary.

### Compute simple two-way interaction

You are free to decide which two variables will form the simple two-way interactions and which variable will act as the third (moderator) variable. In the following R code, we have considered the simple two-way interaction of Referentiality*Gender Status at each level of Anaphor Type

It is recommended to adjust the p-value for multiple testing (Bonferroni correction) by dividing the current $\alpha$-level you declare statistical significance at (i.e., p < 0.05) by the number of simple two-way interaction you are computing (i.e., 2). Thus two-way interaction as statistically significant when $p < 0.025$ (i.e., $p < 0.05/2$). 


```{r}
binary <- filter(prost_2024_singular, Group == "Binary")
nonbinary <- filter(prost_2024_singular, Group == "NonBinary")


two.way.binary <- binary |>
  ezANOVA(dv = diff_score,
          wid = SubjID,
          within = c(Referentiality, Gender_Status),
          within_full = c(Referentiality, Gender_Status, Anteriority),
          type = 3,
          return_aov = F)
kable(two.way.binary$ANOVA)

two.way.nonbinary <- nonbinary |>
  ezANOVA(dv = diff_score,
          wid = SubjID,
          within = c(Referentiality, Gender_Status),
          within_full = c(Referentiality, Gender_Status, Anteriority),
          type = 3,
          return_aov = F)
kable(two.way.nonbinary$ANOVA)
```


### Compute simple simple main effects 

A statistically significant simple two-way interaction can be followed up with simple simple main effects.

Group the data by Group and Gender_Status, and analyze the simple main effect of Referentiality.

#### Effect of Referentiality at each  Gender_Status cell  for Binary Participants
```{r}

ref.effect.G <- binary |>
  filter(Gender_Status == "Gendered") |>
  ezANOVA(dv = diff_score,
          wid = SubjID,
          within = c(Referentiality),
          within_full = c(Referentiality, Gender_Status, Anteriority),
          type = 3,
          return_aov = F)

ref.effect.NG <- binary |>
  filter(Gender_Status == "NonGendered") |>
  ezANOVA(dv = diff_score,
          wid = SubjID,
          within = c(Referentiality),
          within_full = c(Referentiality, Gender_Status, Anteriority),
          type = 3,
          return_aov = F)

kable(ref.effect.G$ANOVA)
kable(ref.effect.NG$ANOVA)
```

#### Effect of Referentiality at each  Gender_Status cell  for Non-Binary Participants
```{r}
ref.effect.G.NB <- nonbinary |>
  filter(Gender_Status == "Gendered") |>
  ezANOVA(dv = diff_score,
          wid = SubjID,
          within = c(Referentiality),
          within_full = c(Referentiality, Gender_Status, Anteriority),
          type = 3,
          return_aov = F)

ref.effect.NG.NB <- nonbinary |>
  filter(Gender_Status == "NonGendered") |>
  ezANOVA(dv = diff_score,
          wid = SubjID,
          within = c(Referentiality),
          within_full = c(Referentiality, Gender_Status, Anteriority),
          type = 3,
          return_aov = F)

kable(ref.effect.G.NB$ANOVA)
kable(ref.effect.NG.NB$ANOVA)
```
### Condition Means 

The N400 effect when antecedents are co-indexed with *himself* or *herself*.  

Significant Effects: **Referentiality; Group X Anteriority; Group x Gender Status x Referentiality**

```{r}

kable(singular_means1 <- prost_2024_singular |> 
  group_by(Group, Referentiality) |>
  summarise(Mean = mean(diff_score), 
            SE = sem(diff_score), 
            SD = sd(diff_score), 
            Max = max(diff_score), 
            Min = min(diff_score)), digits = 2)

kable(singular_means2 <- prost_2024_singular |> 
  filter(Group == "Binary") |>     
  group_by(Group, Gender_Status, Referentiality) |>
  summarise(Mean = mean(diff_score), 
            SE = sem(diff_score), 
            SD = sd(diff_score), 
            Max = max(diff_score), 
            Min = min(diff_score)), digits = 2)

kable(singular_means2 <- prost_2024_singular |> 
  group_by(Group, Anteriority) |>
  summarise(Mean = mean(diff_score), 
            SE = sem(diff_score), 
            SD = sd(diff_score), 
            Max = max(diff_score), 
            Min = min(diff_score)), digits = 2)
```


### Compute simple simple main effects with Bonferroni adjustment using `pwc()` function in the `rstatix`

```{r}
# Pairwise comparisons
pwc <- prost_2024_singular |>
  group_by(Group, Gender_Status) |>
  pairwise_t_test(diff_score ~ Referentiality, paired = TRUE, p.adjust.method = "bonferroni") |>
  select(-p, -.y., -n2 ) # Remove details
kable(pwc)

```

# Visualization: Box plots with p-values 

```{r}

library(ggplot2)
library(ggsci)
library(ggpubr)


bxp <- prost_2024_singular |>
  ggplot(mapping = aes(x = Gender_Status, y = diff_score, colour = Referentiality)) +
  geom_boxplot(staplewidth = .25) +
  facet_wrap(~ Group, ncol = 2) +
  labs(y = "Voltage difference (critical - baseline) in microvolts", x = "Gender Status") +
  theme_classic() +
  scale_color_aaas() +
  scale_fill_aaas(alpha = 0.3)


pwc <- pwc |> add_xy_position(x = "Gender_Status")
bxp + 
  stat_pvalue_manual(pwc, tip.length = 0, hide.ns = TRUE) +
  labs(subtitle = get_test_label(fitted.model.0, detailed = TRUE, correction = "none", row = 11), caption = get_pwc_label(pwc)) +
  xlab("Antecedent Gender Status") +
  labs(colour="Antecedent Referentiality" )
```


## Interaction Plots: Group x Gender Status x Referentiality Interaction
&nbsp;

```{r}
summary_data <- prost_2024_singular |> 
  group_by(Group, Gender_Status, Referentiality) |>
  summarise(mean = mean(diff_score), se = sem(diff_score))|> 
  mutate(lwr = mean - (1.96*se), upr = mean + (1.96*se))
summary_data$label = c("Mary","some woman", "the participant", "someone","Mary","some woman", "the participant", "someone")


ggplot(summary_data, aes(x = Gender_Status, 
                         y = mean,
                         ymin = lwr,
                         ymax = upr,
                         fill = Referentiality, 
                         group = Referentiality,
                         label = label)) + 
  geom_bar(position=position_dodge(), stat="identity", colour = "grey30") + 
  facet_wrap(~ Group, ncol = 2) +
  geom_errorbar(width = .2, colour = "grey30", position = position_dodge(.9))  +
  geom_label_repel(show.legend = FALSE, cex = 2, colour = "black", force = 1) +
  scale_fill_npg() + ylab("Voltage in microvolts") +
  ggtitle("Group by Gender Status by Referentiality Interaction")

```






# Post-hoc tests for Analysis 1: GROUP x ANTERIORITY interaction

The following chunk runs post-hoc tests for the 2-way ***"Group x Anteriority"*** Interaction

```{r echo=TRUE}
# Binary vs Non-Binary Frontal

pander(t.test(diff_score ~ Group
       ,dplyr::filter(prost_2024_singular, (Anteriority == "Frontal"))))

# Binary vs Non-Binary Fronto-Central

pander(t.test(diff_score ~ Group
       ,dplyr::filter(prost_2024_singular, (Anteriority == "FrontoCentral"))))

# Binary vs Non-Binary Central

pander(t.test(diff_score ~ Group
       ,dplyr::filter(prost_2024_singular, (Anteriority == "Central"))))

# Binary vs Non-Binary Centro-Parietal

pander(t.test(diff_score ~ Group
       ,dplyr::filter(prost_2024_singular, (Anteriority == "CentroParietal"))))

# Binary vs Non-Binary Parietal

pander(t.test(diff_score ~ Group
       ,dplyr::filter(prost_2024_singular, (Anteriority == "Parietal"))))
```


# Visualization: Box plots with p-values 

### Compute simple simple main effects with Bonferroni adjustment using `pwc()` function in the `rstatix`

```{r}
# Pairwise comparisons
pwc3 <- prost_2024_singular |>
  group_by(Anteriority) |>
  pairwise_t_test(diff_score ~ Group, p.adjust.method = "bonferroni") |>
  select(-p, -.y. ) # Remove details
kable(pwc3)

```

```{r}
library(ggplot2)
library(ggsci)
library(ggpubr)

bxp3 <- prost_2024_singular |>
  ggplot(mapping = aes(x = Anteriority, y = diff_score, colour = Group)) +
  geom_boxplot(staplewidth = .25) +
  labs(y = "Voltage difference (critical - baseline) in microvolts", x = "Anteriority") +
  theme_classic() +
  scale_color_aaas() +
  scale_fill_aaas(alpha = 0.3)


pwc3 <- pwc3 |> add_xy_position(x = "Anteriority")
bxp3 + 
  stat_pvalue_manual(pwc3, tip.length = 0, hide.ns = TRUE) +
  labs(subtitle = get_test_label(fitted.model.0, detailed = TRUE, correction = "none", row = 7), caption = get_pwc_label(pwc3)) +
  xlab("Anteriority") +
  labs(colour="Group" )
```

##### Interaction Plot: Group x Anteriority

&nbsp;

```{r}
summary_data_3 <- prost_2024_singular |> 
  group_by(Group, Anteriority) |>
  summarise(mean = mean(diff_score), se = sem(diff_score))|> 
  mutate(lwr = mean - (1.96*se), upr = mean + (1.96*se))

ggplot(summary_data_3, aes(x = Anteriority, 
                         y = mean,
                         ymin = lwr,
                         ymax = upr,
                         colour = Group, 
                         linetype= Group, 
                         group = Group)) + 
  geom_point() + 
  geom_line() +
  geom_errorbar(width = .3, position = position_dodge(.1)) +
  scale_colour_npg() + ylab("Voltage in microvolts") +
  ggtitle("Group by Anteriority Interaction")
 # scale_fill_manual(values=c("cadetblue", "lightpink1"))
```


```{r}
summary_data_3 <- prost_2024_singular |> 
  group_by(Group, Anteriority) |>
  summarise(mean = mean(diff_score), se = sem(diff_score))|> 
  mutate(lwr = mean - (1.96*se), upr = mean + (1.96*se))

ggplot(summary_data_3, aes(x = Anteriority, 
                         y = mean,
                         ymin = lwr,
                         ymax = upr,
                         fill = Group,
                         colour = Group,
                         group = Group)) + 
  geom_bar(position=position_dodge(), stat="identity") + 
  geom_errorbar(width = .2, colour = "grey40", position = position_dodge(.9))  +
  scale_fill_npg(alpha = .9) + ylab("Voltage in microvolts") +
  ggtitle("Voltage Differences between groups as a function of Anteriority")
  #scale_fill_manual(values=c("cadetblue", "lightpink1"))
```


